A spatial analysis of the environmental correlates of COVID-19 incidence in Spain

### 2020-04-19 Environmental-Correlates-of-COVID19-Spain/Environmental-Correlates-of-COVID19-Spain.tex
Finished section on selection of variables

The structure of population is a key variable on the incidence. Several
papers related the relationship between a fast trasnmision of COVID-19
and population density (e.g.~cumulative incidence in urban areas like
NYC) and therefore a positive sign will be expected. Note that our
analysis is at provincial level - not a urban level - and therefore is
necessary to rethinking the relationship. In Spain the population is
concentrate (with some exception) in the mediterranean coast (Goerlich
and Cantarino (2013)), in general with high temperature and high
humidity. So, positive or negative sign will be expected in the model.

All those variables will we include in log to capture non-linar
realtionship.

Transit.

@@

\begin{figure}
\centering
\includegraphics{Environmental-Correlates-of-COVID19-Spain_files/figure-latex/weekly-average-incidence-map-1.pdf}
\caption{\label{fig:weekly-average-incidence-map}Mean weekly incidence
of COVID-19 by province, in reported cases by 100,000 people (map omits
Islas Canarias)}
\end{figure}

@@

\begin{verbatim}
## Warning in st_point_on_surface.sfc(sf::st_zm(x)): st_point_on_surface may not
## give correct results for longitude/latitude data

## Warning in st_point_on_surface.sfc(sf::st_zm(x)): st_point_on_surface may not
## give correct results for longitude/latitude data

## Warning in st_point_on_surface.sfc(sf::st_zm(x)): st_point_on_surface may not
## give correct results for longitude/latitude data
\end{verbatim}

\begin{figure}
\centering
\includegraphics{Environmental-Correlates-of-COVID19-Spain_files/figure-latex/map-controls-1.pdf}
\caption{\label{fig:map-controls}Spatial distribution of control
variables by province (maps omits Islas Canarias)}
\end{figure}

@@

\leavevmode\hypertarget{ref-Goerlich2013population}{}%
Goerlich, F.J., Cantarino, I., 2013. A population density grid for
spain. International Journal of Geographical Information Science 27,
2247--2263.

@@

@@

@@

>>>>

The third population variable the we consider is population density,
since it directly affects the contact patterns and contact rates between
individuals in a population (Hu et al., 2013). The evidence available
suggests a positive relationship between the transmission of COVID-19
and population density (e.g.~cumulative incidence in urban areas like
NYC). For this reason, we anticipate a positive relationship between
population density and the incidence of the disease.

The last control variable is the presence of mass transit systems in a
province. Every province in Spain offers some form of public
transportation, however only five provinces have higher order systems of
mass mobility, namely Barcelona, Madrid, Sevilla, Valencia, and Bizkaia.
Public transportation has been hypothesized to relate to the spread of
contagious disease by some researchers using agent-based approaches
(Perez and Dragicevic, 2009; Wang et al., 2010), and while we find scant
evidence of a link in the literature, the idea is intuitively appealing.
After all, unlike the isolation that a car offers to travellers, most
mass transit system are cauldrons of social contact.

@@

\begin{figure}
\centering
\includegraphics{Environmental-Correlates-of-COVID19-Spain_files/figure-latex/weekly-average-incidence-map-1.pdf}
\caption{\label{fig:weekly-average-incidence-map}Mean weekly incidence
of COVID-19 by province, in reported cases by 100,000 people}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics{Environmental-Correlates-of-COVID19-Spain_files/figure-latex/map-controls-1.pdf}
\caption{\label{fig:map-controls}Spatial distribution of control
variables by province}
\end{figure}

@@

@@

\leavevmode\hypertarget{ref-Hu2013scaling}{}%
Hu, H., Nigmatulina, K., Eckhoff, P., 2013. The scaling of contact rates
with population density for the infectious disease models. Mathematical
Biosciences 244, 125--134.
doi:\href{https://doi.org/https://doi.org/10.1016/j.mbs.2013.04.013}{https://doi.org/10.1016/j.mbs.2013.04.013}

@@

\leavevmode\hypertarget{ref-Perez2009agent}{}%
Perez, L., Dragicevic, S., 2009. An agent-based approach for modeling
dynamics of contagious disease spread. International journal of health
geographics 8, 50.

@@

\leavevmode\hypertarget{ref-Wang2010gis}{}%
Wang, J., Xiong, J., Yang, K., Peng, S., Xu, Q., 2010. Use of gis and
agent-based modeling to simulate the spread of influenza, in: 2010 18th
International Conference on Geoinformatics. IEEE, pp. 1--6.
